How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-06-14 Paper/paper.tex
added section Insights for Computing Education

\subsubsection{Exact clones between sprites}
With occurrences in 11\% of the Scratch projects in our dataset, the use of exactly identical clones between sprites is relatively common. In a sense, the Scratch users are not to blame here, as Scratch does not support procedure calls between sprites, only within them. So in many cases there is no way to share the functionality other than by making a copy. We are not aware of the underlying rationale of the Scratch team that lead to this decision, however it seems that a large part of the Scratch users would use the functionality to call procedures between sprites.

@@

>>>>

\subsubsection{Exact clones between sprites}
\label{implicationsClones}
With occurrences in 11\% of the Scratch projects in our dataset, the use of exactly identical clones between sprites is relatively common. In a sense, the Scratch users are not to blame here, as Scratch does not support procedure calls between sprites, only within them. So in many cases there is no way to share the functionality other than by making a copy. We are not aware of the underlying rationale of the Scratch team that lead to this decision, however it seems that a large part of the Scratch users would use the functionality to call procedures between sprites.

@@

\subsection{Insights for computing education}
Our findings confirm that Scratch is mainly used for its intended purpose as a first-exposure programming environment for creating simple programs and interactive animations: the majority of programs are small and more than half are interactive and contain no conditional statements.
The analysis also indicates that computational concepts like conditionals and variables are being applied ---conditional statements are found in 40\% of the projects and variables in 32\%. The same does not hold for loops: even though 77\% of the projects contain loops, only 14\% contains conditional loops. We attribute the increased use of the \texttt{forever} loop to the Scratch language design and we are skeptical over whether it indicates an understanding of loop concepts.
Only 8\% of the projects contain procedures, and those use them a lot and for more complex code, which is an indication of use by more experienced programmers. This very essential programming concept is therefore not sufficiently applied, which can be attributed both to limitations imposed by the Scratch programming environment, like the local scope of procedures, and to the difficulty for internalizing certain computational thinking concepts before a certain age \cite{seiter_2013}.

The high occurrence of cloned scripts (in 26\% of the projects) could be the result of the limited use of procedures. Other code smells are also frequently found: 28\% of the projects include dead code and 30\% have large scripts. Knowing from prior research that long scripts and code duplication decreases ability to understand and modify Scratch programs \cite{hermans_2016}, and that Scratch programmers tend to exhibit certain bad programming habits \cite{Meerbaum_habits_2011}, we believe educating novice programmers on code quality concepts is an issue that should be further researched.

It must be noted that the scope of this study includes the programming artifacts alone. Our findings are limited by the lack of (1) process data and (2) data on the age and other characteristics related to the programmers. However, the project data that is available in our dataset can facilitate further studies on computing education, and this is the reason that we are publishing it.
